Deployable Probabilistic Programming

### 2019-04-29 paper/paper.tex
a touch
convinced

In our journey towards implementation and deployment of Infergo, we
considered options and made implementation decisions which we
believe to have been crucial for addressing the challenges and
following the guidelines we formulated for implementation of a
deployable probabilistic programming facility. While different
choices may be as well suited for the purpose as the ones we
made, we feel important to share our contemplations and justify
decisions, to which the rest of this section is dedicated, from
choosing the implementation language, through details of
automatic differentiation, to support for inference on streaming
and stochastic observations.

@@

// Simulate simulates y based on x and 
// regression parameters.
func (m *LRModel) Simulate(x, alpha, beta float64)
    float64 {
  y := alpha + beta*x
  return y
}
\end{lstlisting}
\caption{Linear regression: the simulator is re-used in both
  inference and prediction.}
\label{fig:linear-regression}
\end{figure}
If only the maximum \textit{a posteriori} is estimated, the
parameter values will maximize their likelihood given the
observations:
\begin{equation}
	(\alpha, \beta, \sigma)_{MAP} = \arg \max \prod\limits_i \mathrm{Normal}(\alpha + \beta x_i, \sigma)
	\label{eqn:linear-regression-2}
\end{equation}

@@

We use three models of different size and structure for the
comparison: two of the models, Eight schools and Gaussian
mixture model, were introduced as a part of the case
studies~(Section~\ref{sec:case-studies}); the code for the third
model, Latent Dirichlet allocation~\cite{BNG03}, is included in
Appendix~\ref{app:lda}. The code and data for all models were
originally included in Stan examples and translated for use
with Turing and Infergo. \textit{Eight schools} is the smallest model.
One may expect that most of the time execution time is spent in
executing the boilerplate code rather than the model. \textit{Gaussian
mixture model} is a more elaborated model which benefits greatly
from vectorization. The greater the number of observations (1000
data points were used in the evaluation), the more significant
is the benefit of vectorized computations. \textit{Latent
Dirichlet allocation} is, again, a relatively elaborated model,
that is, however, not as easy to vectorize.

>>>>

In our journey towards implementation and deployment of Infergo, we
considered options and made implementation decisions which we
believe to have been crucial for addressing the challenges and
following the guidelines we formulated for implementation of a
deployable probabilistic programming facility. While different
choices may be as well suited for the purpose as the ones we
made, we feel it is important to share our contemplations and
justify decisions, to which the rest of this section is
dedicated, from choosing the implementation language, through
details of automatic differentiation, to support for inference
on streaming and stochastic observations.

@@

// Simulate simulates y based on x and 
// regression parameters.
func (m *LRModel) Simulate(x, alpha, beta float64)
    float64 {
  y := alpha + beta*x
  return y
}
\end{lstlisting}
\caption{Linear regression: the simulator is re-used in both
  inference and prediction.}
\label{fig:linear-regression}
\end{figure}
If only the maximum \textit{a posteriori} is estimated, the
parameter values will maximize their likelihood given the
observations:
\begin{equation}
	(\alpha, \beta, \sigma)_{MAP} = \arg \max \prod\limits_i \mathrm{pdf}_\mathrm{Normal}(\alpha + \beta x_i, \sigma)
	\label{eqn:linear-regression-2}
\end{equation}

@@

We use three models of different size and structure for the
comparison: two of the models, Eight schools and Gaussian
mixture model, were introduced as a part of the case
studies~(Section~\ref{sec:case-studies}); the code for the third
model, Latent Dirichlet allocation~\cite{BNG03}, is included in
Appendix~\ref{app:lda}. The code and data for all models were
originally included in Stan examples and translated for use
with Turing and Infergo. \textit{Eight schools} is the smallest model.
One may expect that most of the time execution time is spent in
executing the boilerplate code rather than the model. \textit{Gaussian
mixture model} is a more elaborated model which benefits greatly
from vectorization. The greater the number of observations (1000
data points were used in the evaluation), the more significant
is the benefit of vectorized computations. \textit{Latent
Dirichlet allocation} is, again, a relatively elaborated model,
which is, however, not as easy to vectorize.
